## Prediction of Exercise Manner from Health Device Data

# Executive Summary

I used a tree classification model and a linear disriminant model to predict the classe outcome. The tree classification model was accurate 40% of the time on the testing data. On the values for which the tree classification model was incorrect, the linear discriminant model was correct 50% of the time. Therefore, in combination, the 2 models where correct 70% of the time (14 / 20). 

Several additional models were attempted, including a random forest and naive bayes, however the time to compute each was too long to determine whether they were good models.

# Setup

Set working directory

```{r,echo=TRUE}
setwd("~/Training/Coursera/Machine Learning")
```

Load the training and testing data

```{r,echo=TRUE}
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```

Load R packages

```{r,echo=TRUE}
library(ggplot2)
library(plyr)
library(stats)
library(caret)
library(rattle)
```

# Refine the data

Eliminate columns with greater than 50% N/A values

```{r,echo=TRUE}
training2 <- training[colSums(!is.na(training))>nrow(training)*0.5]
```

Eliminate columns which have close to 0 variability

```{r,echo=TRUE}
nsv <- nearZeroVar(training2,saveMetrics=TRUE)
training3 <- training2[-c(6,12:20,43:48,52:60,74:82)]
```

Eliminate the variable x # x just denotes the row number

```{r,echo=TRUE}
training4 <- training3[-c(1)]
```

Remove the timestamp data

```{r,echo=TRUE}
training5 <- training4[-c(2:4)]
```

# Build a classification tree model

```{r,echo=TRUE}
modFit <- train(classe~.,method="rpart",data=training5)
print(modFit$finalModel)
fancyRpartPlot(modFit$finalModel)
```

# Build a linear discriminant model

```{r,echo=TRUE}
modFit2 <- train(classe~.,method="lda",data=training5)
```


# Predict the values of the testing data
```{r,echo=TRUE}
pred <- predict(modFit,testing)
pred2 <- predict(modFit2,testing)
```

The values predicted by the classification tree model are `r pred`
The values predicted by the linear discriminant model are `r pred2`